Church Bells are Not Ringing: Decrease in the Opposite-Sex and Same-Sex Couples Opting for Religious Marriage Ceremonies in England and Wales

### 2024-04-25 paper/paper.qmd
final paper upload

---
title: "[WIP] Church Bells are Not Ringing: Decrease in the Opposite-Sex and Same-Sex Couples Opting for Religous Marriage Ceremonies in England and Wales"
subtitle: "Analysis and Comparative Study of Marriage Ceremony Type (Civil and Religious) for Opposite-Sex (1970 to 2020) and Same-Sex (2016 to 2020) Couples "
author: 
  - Francesca Ye
thanks: "Code and data are available at: https://github.com/Francesca-Ye/marriage_ceremonies_religion"
date: today
date-format: long
abstract: "This paper investigates the annual number of marriage conducted in England and Wales based on ceremony type recorded the by the Office of National Statistics. Based on the data set, these trends between 1970 to 2020 were analyzed for opposite-sex couples and between 2016 to 2020 for same-sex couples in tandem with comparative study on couple type preferences. Analysis found that civil ceremonies have overtake civil cermonies as the predominant ceremony type for both couple types. The results of this paper indicate that religion is less of an inetgral facet of life in England and Wales as indicated by religion becoming less synonymous with marriage."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---
```{r}
#| message: false
#| echo: false
#| warning: false
#### Work space set-up ####

@@

# Introduction {#sec-intro}
In the United Kingdom, there has been increasing research and attention towards the decline of the importance of religion in individual life. The Policy Institute at King’s College London has collected data which corroborates this trend. In 1981, 57% of Britons identified as religious while that number has decreased to 32% in 2022. Similarly, 75% of Britons believed in God when surveyed in 1981 and that number has significantly fallen to 49% in 2022 [@ukreligion]. However, these trends are not unique to the United Kingdom, religiosity across most Western countries has declined over the past couple of decades. The number of people in the United States that identified as religious fell from 82% in 1982 to 58% in 2017 [@ukreligion].

@@

This paper takes data from the *Office for National Statistics in the United Kingdom* from the 2020 edition of the "Marriages in England and Wales" data set to examine if the presence of religion in legal marriage has declined over time in tandem with answering if the result is difference between same and opposite sex couples. The estimand of interest is the presence of religion in consolidating a legal marriage. To examine whether or not religion in marriage has declined, graphical analysis was conducted to visualize the annual data on marriage ceremony type between 1970 to 2020 for opposite-sex couples and 2016 to 2020 for same-sex couples. Summary statistics were also calculated to get a greater idea of the numerical significance of religion in this space.

This paper finds that religion in the context of marriage has declined over the years in addition to same-sex couples more strongly preferring civil ceremonies in comparison to opposite-sex couples. More specifically, religious ceremonies appears to be rapidly declining in popularity amongst opposite-sex couples over the past five decades whilst civil ceremonies appear to be comparatively more stable in annual totals. This could imply that religiosity has decreased in England and Wales as couples are choosing to not involve religion in a significant life event. Interestingly, same-sex couples have strongly preferred civil ceremonies since the legalization of same-sex marriage in 2014 which makes a difference between opposite-sex couples. The latter has seen an active decline whilst the former has maintained an aversion to religious marriage ceremonies. Although both result in a fewer religious ceremonies, this indicates that the move away from religion could be motivated by different factors such as homophobia within religion for same-sex couples.

@@

# Data {#sec-data}
## Data Description
The data sets for same and opposite-sex couples were derived from the *Office for National Statistics in the United Kingdom* with the 2020 edition of the “Marriages in England and Wales” workbook as raw data. This data is collected at the time a marriage is registered as part of civil registration which is a legal requirement in completing a legally recognized marriage in England and Wales. Annual marriage statistics are available to the public approximately two years after the end of the reference year [@statsguide]. Notably, same-sex couples that are currently in a civil partnership can have it converted to a marriage but these numbers are not counted towards the marriage statistics of same-sex couples in the workbook. The data collected and used in this paper is used to measure the number of couples that opt for civil or religious marriage ceremonies on a yearly basis [@statsguide].

By measuring marriage ceremony trends, it can help measure and provide insight onto greater social attitudes towards religion. Based on aforementioned literature about the perceived decline of religiosity in the United Kingdom, marriage ceremony data was chosen as a potential form of measuring this phenomena due to the historical link between marriage and religion [@marriagedeclineuk]. Since marriage is a significant life event, the presence and absence of religion in ceremonies can corroborate the decline of religiosity in individual life. These marriage ceremony entries in the "Marriages in England and Wales" workbook are a reflection of each couple's relationship with religion and whether or not it is strong enough to opt for a religious ceremony. Although the data set only reflects ceremony types, those entries were antedated by personal consideration of a couple's proximity to religion.

@@

# Results {#sec-results}
Looking at the graphed data and summary statistics, there appears to be a clear decline in the number of opposite-sex couples opting for a religious ceremony across a fifty-year time frame between 1970 and 2020. For same-sex couples, civil ceremonies have consistently and overwhelmingly been the most popular ceremony type since the legalization of same-sex marriage in 2014.Looking further into the data sets, it is apparent that preliminary conclusions from the summary data in @sec-data hold true after further analysis.

@@

While @fig-oppositecivilvsreligious shows that the annual total of religious ceremonies as a whole have decreased, the annual total of of religious ceremonies sub categorized by type reveal a similar story. @fig-oppositereligioustype highlights that all Christian denominations have seen a decrease in the annual total of ceremonies conducted. While all other religions appears to be an outlier in this instance, the number of non-Christian religious ceremonies conducted has consistently been the least conducted type of ceremony amongst opposite-sex couples. Additionally, since the collection of data on non-Christian religious ceremonies, the standard deviation is relatively small at 810.76 which further highlights the minimal impact non-Christian ceremonies have on the larger data set combined with their unpopularity. As the only category with growth for religious ceremonies, this could mean that religiosity is not growing immensely outside of religiosity.

It is also apparent that Christianity has decreased in popularity for marriage ceremonies. Comparatively, Roman Catholic and other Christian denominations have seen decreases over a fifty-year period to a significant extent. The standard deviation for Roman Catholic is 10,788.53 and 11,781.59 for other Christian denominations which is considerably varied and highlights the significant drop in marriage ceremonies of these sects as seen in graphical analysis. However, the most notable drop would be Church of England ceremonies which had a standard deviation of 38,034.88, the largest, and when factoring in the downward graphical trend, it is clear that the Church of England has experienced major decreases in popularity for marriage ceremonies. Resultantly, Christian religious ceremonies have experienced somewhat extreme decreases in popularity that could imply changing individual and social attitudes on religion.

@@

By looking at the most recent decade, the decline in religious ceremonies still appears to be rapid. In @fig-oppositereligioustypeten, the peak of the decade for Church of England ceremonies is 2012 with 79,473; however, the lowest year excluding 2020 is 2019 with 39,897 ceremonies. There is almost a 40,000 ceremony drop between these two years which further indicate the decline of religious ceremonies and rapid pacing just over a ten-year period. As the most popular religion for marriage ceremonies by a considerable margin, the Church of England still appears to be experiencing significant drops in popularity.

@@

@tbl-oppositedecades shows the total number of religious and civil ceremonies across five decades between 1971 and 2020. In the 1970s, religious ceremonies are more popular by just under 200,000 more ceremonies conducted. While both ceremony types see a decrease in the 1980s, the gap between the two is only about 100,000 with religious ceremonies still leading. Interestingly, civil ceremonies appear to minimally fluctuate by 100,000 ceremonies over the next three decades whilst religious ceremonies actively decrease in their decade summary. Religious ceremonies conducted in the 1970s was 1,995,792 in comparison to 567,934 in the 2010s. This is contrasted by civil ceremonies which saw 1,823,284 and 1,674,720 in the very same decades. Considering the decade summaries, it is clear that civil ceremonies have not necessarily increased by taking away from couples that might have opted for religious ceremonies due to their relative stability in numbers over time. It can also be ascertained that religious ceremonies have experienced a severe decreased over the past five decades as the annual total of ceremonies being conducted presently is just over a quarter to the number of ceremonies in the 1970s.

@@

As seen in @fig-samecivilvsreligious, civil ceremonies for same-sex couples have consistently been the premier ceremony choice for couples intending to marry. @fig-samesexdemographics visualizes the annual number of couples that marry by sex. As shown in @fig-samesexdemographics, female couples have consistently married more than male couples in any given year. However, the gap between these totals are not large with 2019 representing the largest gap of 996 marriages. It can be derived that sex does not play a significant factor in the popularity of civil ceremonies as females do not make up a large majority of all same-sex married couples. Therefore, the unpopularity of religious ceremonies could be due to other factors which is touched upon in @sec-discussion.

@@

As same-sex marriage was only recently legalized, the six years of data was summed up by civil and religious to provide a greater insight on the preferences of same-sex couples. @tbl-samesextotals provides the total number of religious and civil ceremonies conducted for same-sex couples since 2014 and up to 2020. Civil ceremonies are clearly the more popular choice with just under 30,000 more ceremonies conducted. Of all same-sex marriage ceremonies conducted since the the inception of legalization, religious ceremonies only make up about 0.76% whilst civil ceremonies make up the rest. This differs with opposite-sex marriages as they have experienced a decline of religious ceremonies whereas religious ceremonies for same-sex marriages have been extremely unpopular from the outset of legalization.

@@

Political science professor at the University of Michigan, Ronald F. Inglehart, wrote in his book *Religion's Sudden Decline: What's Causing it, and What Comes Next?* on the decline of religiosity. Inglehart's main argument on the reason for the decline in religion across the world is due to the historical stance of most major religions advising women to have as many children possible due to high infant-mortality rates. He argues that the introduction of individual-choice norms such as the normalization of divorce, acceptance of abortion, gender equality, and acceptance of LGBTQIA+ identities has empowered individuals to make decisions outside of religious guidance [@inglehart]. Connecting this with the results gathered in this paper, it is apparent that the decline of religious marriages could specifically due to changing social norms outside of religion. As the societal pressure to reproduce and have children solely after marriage decreases, it makes sense to assume religious marriage ceremonies would decrease. This follows Inglehart's argument that the change towards individual-choice norms has undermined specific directives and the overall power of religions against societies. Extrapolating further, the dilution of religion in individual life could also result in its removal from significant life events such as marriage which was seen in analysis of marriage ceremony data in England and Wales.

## Implication: Fewer Citizens Identify as Christian
Interestingly, the marriage ceremony data analysed and discussed in this paper also indicate that Christianity is becoming less popular or prominent in the lives of English and Welsh citizens. As seen in @fig-oppositereligioustypes, all sects of Christianity have experienced a decrease in the annual number of marriage ceremonies conducted. This implies that less couples are religiously involved in Christianity to the extent of including it in their marriage plans.

Looking to "Religion, England and Wales: Census 2021" from the *Office for National Statistics* provides further insight on whether or not religious identification and marriage ceremony type are congruent in their data. In the 2021 census, 46.2% (27.5 million) identified as Christian in comparison to 59.3% (33.3 million) in 2011 [@census2021]. Although Christianity still remains the most popular religion in England and Wales, the decrease over a decade is still somewhat significant as less than half the population now identifies as Christian. This reflected on a smaller level in the opposite-sex marriage ceremony data, as seen in @fig-oppositereligiousten, since all Christian sects experienced a decline in annual marriage ceremonies with the Church of England seeing a reduction of just under 40,000 ceremonies in 2020 compared to 2011.

@@

## Implication: Same-Sex Couples Overwhelmingly Prefer Civil Ceremonies
As previously discussed, same-sex couples differ to opposite-sex couples in regards to their marriage ceremony decision-making behaviour. Based on the data examined in this paper, same-sex couples have consistently and overwhelmingly opted for civil ceremonies over religious ceremonies since the legalization of same-sex marriage in 2014. Comparing marriage ceremony type of same-sex couples in @fig-samecivilvsreligious and opposite-sex couples in @fig-oppositecivilvsreligious, the trend lines are vastly different. For same-sex couples, civil ceremonies make up over 99% of all annual marriages conducted whereas opposite-sex couples have experienced periods where religious ceremonies were the predominant type and still have a significant presence in accounting for marriage ceremony types.

@@

## Limitations
Despite this paper's effort to highlight how marriage ceremony data can be a reflection of the religious attitudes in England and Wales, there are some limitations that must be addressed.

@@

There are also accuracy considerations when it comes to making conclusions based on this data. As noted by the *Office for National Statistics*, data is taken directly from marriage registries from across the two countries and the information is mainly supplied by the people getting married without any major statistical verification aside from edits for cleaning. Couples that get married abroad are also not recorded with the exceptions being consular and military marriages [@marriageqmi]. Marriage ceremony by type could also be inherently misleading as religious marriages, such as Muslim or Sikh, can take place at unregistered premises [@marriageqmi]. The legal formalization of these marriages often occur as civil ceremonies and are recorded as such without an ability to account for the religious aspect of their marriage plans. As a result, the annual number of civil marriages could possibly inflated with the inclusion of couples that have separate legal and religious marriage ceremonies.

There could also be ethical implications to using this data as a reflection of religiosity. Using the same-sex couple data as an example, due to the legalization of same-sex marriage only occurring in 2014 it can be argued that there is not enough data to make conclusions about ceremony preferences. As well, it could be problematic to assume same-sex couples opt for civil ceremonies due to homophobia in religion and therefore are not religious. It is possible that same-sex couples could be practicing religion outside of mainstream institutions and the preference for civil ceremonies is due to these institutions not approving ceremonies. 
Another possible limitation is the financial costs associated with a religious versus civil ceremony. The average cost at minimum for a wedding in the Church of England is £539 [@churchcost] whilst the cheapest civil ceremony is £46 [@registercost]. Due to the price difference in religious ceremonies, couples might also opt for a civil ceremony to save on costs and therefore the motivation would not be an aversion to religion.

## Future Research {#sec-future}
Beyond this paper, there is a lot of research that could be undertaken to further understand the presence of religion and how it influences different facets of life. Looking to marriage ceremonies specifically, more research could be done on the factors influencing why couples opt for a civil or religious ceremony. Better data collection methods could also be developed to better collect data on marriage solemnization outside of the legal form in order to gain insight on the prevalence of religious ceremonies that are conducted outside of legally approved premises. By collecting a greater depth of marriage ceremony data, future research will be able to better ascertain or disprove whether or not marriage ceremony data is a good reflection of religiosity in a country.

There are also other areas that could be examined as an indicator of a country or population's religiosity such as enrollment in religious versus secular schools, support or use of abortion services, distribution of charitable donations, and observance of religious holidays on mandated holidays for that purpose. Delving further into research on charitable donations, examining the annual charitable donations of a country and analysing the amounts that go to each type of charity would be a novel way to see how religiosity of lack thereof can impact donation decisions.

>>>>

---
title: "Church Bells are Not Ringing: Decrease in the Opposite-Sex and Same-Sex Couples Opting for Religious Marriage Ceremonies in England and Wales"
subtitle: "Analysis and Comparative Study of Marriage Ceremony Type (Civil and Religious) for Opposite-Sex (1970 to 2020) and Same-Sex (2016 to 2020) Couples "
author: 
  - Francesca Ye
thanks: "Code and data are available at: https://github.com/Francesca-Ye/marriage_ceremonies_religion"
date: today
date-format: long
abstract: "This paper investigates the annual number of marriages conducted in England and Wales based on ceremony type recorded the by the Office of National Statistics. Based on the data set, these trends between 1970 to 2020 were analyzed for opposite-sex couples and between 2016 to 2020 for same-sex couples in tandem with a comparative study on couple type preferences. Analysis found that civil ceremonies have overtaken civil ceremonies as the predominant ceremony type for both couple types. The results of this paper indicate that religion is less of an integral facet of life in England and Wales as indicated by religion becoming less synonymous with marriage."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---
```{r}
#| message: false
#| echo: false
#| warning: false
#### Work space set-up ####

@@

# Introduction {#sec-intro}
In the United Kingdom, there has been increasing research and attention towards the decline of the importance of religion in individual life. The Policy Institute at King’s College London has collected data which corroborates this trend. In 1981, 57% of Britons identified as religious while that number has decreased to 32% in 2022. Similarly, 75% of Britons believed in God when surveyed in 1981 and that number has significantly fallen to 49% in 2022 [@ukreligion]. However, these trends are not unique to the United Kingdom, religiosity across most Western countries has declined over the past couple of decades. The number of people in the United States who identified as religious fell from 82% in 1982 to 58% in 2017 [@ukreligion].

@@

This paper takes data from the *Office for National Statistics in the United Kingdom* from the 2020 edition of the "Marriages in England and Wales" data set to examine if the presence of religion in legal marriage has declined over time in tandem with answering if the result is difference between same and opposite-sex couples. The estimand of interest is the presence of religion in consolidating a legal marriage. To examine whether or not religion in marriage has declined, graphical analysis was conducted to visualize the annual data on marriage ceremony type between 1970 to 2020 for opposite-sex couples and 2016 to 2020 for same-sex couples. Summary statistics were also calculated to get a greater idea of the numerical significance of religion in this space.

This paper finds that religion in the context of marriage has declined over the years in addition to same-sex couples more strongly preferring civil ceremonies in comparison to opposite-sex couples. More specifically, religious ceremonies appear to be rapidly declining in popularity amongst opposite-sex couples over the past five decades whilst civil ceremonies appear to be comparatively more stable in annual totals. This could imply that religiosity has decreased in England and Wales as couples are choosing to not involve religion in a significant life event. Interestingly, same-sex couples have strongly preferred civil ceremonies since the legalization of same-sex marriage in 2014 which makes a difference between opposite-sex couples. The latter has seen an active decline whilst the former has maintained an aversion to religious marriage ceremonies. Although both result in fewer religious ceremonies, this indicates that the move away from religion could be motivated by different factors such as homophobia within religion for same-sex couples. The findings of this paper could be indicators of an overall decline in religiosity in England and Wales as marriage ceremonies could potentially be a direct reflection of individual religious beliefs.

@@

# Data {#sec-data}
## Data Description
The data sets for same and opposite-sex couples were derived from the *Office for National Statistics in the United Kingdom* with the 2020 edition of the “Marriages in England and Wales” workbook as raw data. This data is collected at the time a marriage is registered as part of civil registration which is a legal requirement in completing a legally recognized marriage in England and Wales. Annual marriage statistics are available to the public approximately two years after the end of the reference year [@statsguide]. Notably, same-sex couples that are currently in a civil partnership can have it converted to a marriage but these numbers are not counted towards the marriage statistics of same-sex couples in the workbook. The data collected and used in this paper is used to measure the number of couples that opt for civil or religious marriage ceremonies on an annual basis [@statsguide].

Measuring marriage ceremony trends, can help measure and provide insight into greater social attitudes towards religion. Based on the aforementioned literature about the perceived decline of religiosity in the United Kingdom, marriage ceremony data was chosen as a potential form of measuring this phenomenon due to the historical link between marriage and religion [@marriagedeclineuk]. Since marriage is a significant life event, the presence and absence of religion in ceremonies can corroborate the decline of religiosity in individual life. These marriage ceremony entries in the "Marriages in England and Wales" workbook are a reflection of each couple's relationship with religion and whether or not it is strong enough to opt for a religious ceremony. Although the data set only reflects ceremony types, those entries were antedated by personal consideration of a couple's proximity to religion.

@@

# Results {#sec-results}
Looking at the graphed data and summary statistics, there appears to be a clear decline in the number of opposite-sex couples opting for a religious ceremony across a fifty-year time frame between 1970 and 2020. For same-sex couples, civil ceremonies have consistently and overwhelmingly been the most popular ceremony type since the legalization of same-sex marriage in 2014. Looking further into the data sets, it is apparent that preliminary conclusions from the summary data in @sec-data hold after further analysis.

@@

While @fig-oppositecivilvsreligious shows that the annual total of religious ceremonies has decreased, the annual total of religious ceremonies sub-categorized by type reveal a similar story. @fig-oppositereligioustype highlights that all Christian denominations have seen a decrease in the annual total of ceremonies conducted. While all other religions appear to be an outlier in this instance, the number of non-Christian religious ceremonies conducted has consistently been the least conducted type of ceremony amongst opposite-sex couples. Additionally, since the collection of data on non-Christian religious ceremonies, the standard deviation is relatively small at 810.76 which further highlights the minimal impact non-Christian ceremonies have on the larger data set combined with their unpopularity. As the only category with growth for religious ceremonies, this could mean that religiosity is not growing immensely outside of religiosity.

It is also apparent that Christianity has decreased in popularity for marriage ceremonies. Comparatively, Roman Catholic and other Christian denominations have seen decreases over a fifty-year period to a significant extent. The standard deviation for Roman Catholic is 10,788.53 and 11,781.59 for other Christian denominations which are considerably varied and highlight the significant drop in marriage ceremonies of these sects as seen in graphical analysis. However, the most notable drop would be Church of England ceremonies which had a standard deviation of 38,034.88, the largest, and when factoring in the downward graphical trend, it is clear that the Church of England has experienced major decreases in popularity for marriage ceremonies. Resultantly, Christian religious ceremonies have experienced somewhat extreme decreases in popularity that could imply changing individual and social attitudes toward religion.

@@

By looking at the most recent decade, the decline in religious ceremonies still appears to be rapid. In @fig-oppositereligioustypeten, the peak of the decade for Church of England ceremonies is 2012 with 79,473; however, the lowest year excluding 2020 is 2019 with 39,897 ceremonies. There is almost a 40,000 ceremony drop between these two years which further indicates the decline of religious ceremonies and rapid pacing just over a ten-year period. As the most popular religion for marriage ceremonies by a considerable margin, the Church of England still appears to be experiencing significant drops in popularity.

@@

@tbl-oppositedecades shows the total number of religious and civil ceremonies across five decades between 1971 and 2020. In the 1970s, religious ceremonies were more popular with just under 200,000 more ceremonies conducted. While both ceremony types saw a decrease in the 1980s, the gap between the two is only about 100,000 with religious ceremonies still leading. Interestingly, civil ceremonies appear to minimally fluctuate by 100,000 ceremonies over the next three decades whilst religious ceremonies actively decrease in their decade summary. Religious ceremonies conducted in the 1970s were 1,995,792 in comparison to 567,934 in the 2010s. This is contrasted by civil ceremonies which saw 1,823,284 and 1,674,720 in the same decades. Considering the decade summaries, it is clear that civil ceremonies have not necessarily increased by taking away from couples that might have opted for religious ceremonies due to their relative stability in numbers over time. It can also be ascertained that religious ceremonies have experienced a severe decrease over the past five decades as the annual total of ceremonies being conducted presently is just over a quarter of the number of ceremonies in the 1970s.

@@

As seen in @fig-samecivilvsreligious, civil ceremonies for same-sex couples have consistently been the premier ceremony choice for couples intending to marry. @fig-samesexdemographics visualizes the annual number of couples that marry by sex. As shown in @fig-samesexdemographics, female couples have consistently married more than male couples in any given year. However, the gap between these totals is not large with 2019 representing the largest gap of 996 marriages. It can be derived that sex does not play a significant factor in the popularity of civil ceremonies as females do not make up a large majority of all same-sex married couples. Therefore, the unpopularity of religious ceremonies could be due to other factors which are touched upon in @sec-discussion.

@@

As same-sex marriage was only recently legalized, the six years of data was summed up by civil and religious to provide a greater insight into the preferences of same-sex couples. @tbl-samesextotals provides the total number of religious and civil ceremonies conducted for same-sex couples since 2014 and up to 2020. Civil ceremonies are clearly the more popular choice with just under 30,000 more ceremonies conducted. Of all same-sex marriage ceremonies conducted since the the inception of legalization, religious ceremonies only make up about 0.76% whilst civil ceremonies make up the rest. This differs with opposite-sex marriages as they have experienced a decline in religious ceremonies whereas religious ceremonies for same-sex marriages have been extremely unpopular from the outset of legalization.

@@

Political science professor at the University of Michigan, Ronald F. Inglehart, wrote in his book *Religion's Sudden Decline: What's Causing it, and What Comes Next?* on the decline of religiosity. Inglehart's main argument on the reason for the decline in religion across the world is due to the historical stance of most major religions advising women to have as many as children possible due to high infant mortality rates. He argues that the introduction of individual-choice norms such as the normalization of divorce, acceptance of abortion, gender equality, and acceptance of LGBTQIA+ identities has empowered individuals to make decisions outside of religious guidance [@inglehart]. Connecting this with the results gathered in this paper, it is apparent that the decline of religious marriages could be specifically due to changing social norms outside of religion. As the societal pressure to reproduce and have children solely after marriage decreases, it makes sense to assume religious marriage ceremonies would decrease. This follows Inglehart's argument that the change towards individual-choice norms has undermined specific directives and the overall power of religions against societies. Extrapolating further, the dilution of religion in individual life could also result in its removal from significant life events such as marriage which was seen in analysis of marriage ceremony data in England and Wales.

## Implication: Fewer Citizens Identify as Christian
Interestingly, the marriage ceremony data analysed and discussed in this paper also indicate that Christianity is becoming less popular or prominent in the lives of English and Welsh citizens. As seen in @fig-oppositereligioustypes, all sects of Christianity have experienced a decrease in the annual number of marriage ceremonies conducted. This implies that fewer couples are religiously involved in Christianity to the extent of including it in their marriage plans.

Looking at "Religion, England and Wales: Census 2021" from the *Office for National Statistics* provides further insight into whether or not religious identification and marriage ceremony type are congruent in their data. In the 2021 census, 46.2% (27.5 million) identified as Christian in comparison to 59.3% (33.3 million) in 2011 [@census2021]. Although Christianity remains the most popular religion in England and Wales, the decrease over a decade is still somewhat significant as less than half the population now identifies as Christian. This is reflected on a smaller level in the opposite-sex marriage ceremony data, as seen in @fig-oppositereligiousten, since all Christian sects experienced a decline in annual marriage ceremonies with the Church of England seeing a reduction of just under 40,000 ceremonies in 2020 compared to 2011.

@@

## Implication: Same-Sex Couples Overwhelmingly Prefer Civil Ceremonies
As previously discussed, same-sex couples differ from opposite-sex couples in regards to their marriage ceremony decision-making behaviour. Based on the data examined in this paper, same-sex couples have consistently and overwhelmingly opted for civil ceremonies over religious ceremonies since the legalization of same-sex marriage in 2014. Comparing marriage ceremony types of same-sex couples in @fig-samecivilvsreligious and opposite-sex couples in @fig-oppositecivilvsreligious, the trend lines are vastly different. For same-sex couples, civil ceremonies make up over 99% of all annual marriages conducted whereas opposite-sex couples have experienced periods where religious ceremonies were the predominant type and still have a significant presence in accounting for marriage ceremony types.

@@

## Limitations
Despite this paper's effort to highlight how marriage ceremony data can be a reflection of the religious attitudes in England and Wales, some limitations must be addressed.

@@

There are also accuracy considerations when it comes to making conclusions based on this data. As noted by the *Office for National Statistics*, data is taken directly from marriage registries from across the two countries and the information is mainly supplied by the people getting married without any major statistical verification aside from edits for cleaning. Couples that get married abroad are also not recorded with the exceptions being consular and military marriages [@marriageqmi]. Marriage ceremony by type could also be inherently misleading as religious marriages, such as Muslim or Sikh, can take place at unregistered premises [@marriageqmi]. The legal formalization of these marriages often occur as civil ceremonies and are recorded as such without an ability to account for the religious aspect of their marriage plans. As a result, the annual number of civil marriages could be inflated with the inclusion of couples that have separate legal and religious marriage ceremonies.

There could also be ethical implications to using this data as a reflection of religiosity. Using the same-sex couple data as an example, due to the legalization of same-sex marriage only occurring in 2014 it can be argued that there is not enough data to make conclusions about ceremony preferences. It could also be problematic to assume same-sex couples opt for civil ceremonies due to homophobia in religion and therefore are not religious. It is possible that same-sex couples could be practicing religion outside of mainstream institutions and the preference for civil ceremonies is due to these institutions not approving ceremonies. 
Another possible limitation is the financial costs associated with a religious versus civil ceremony. The average cost, at minimum, for a wedding in the Church of England is £539 [@churchcost] whilst the cheapest civil ceremony is £46 [@registercost]. Due to the price difference in religious ceremonies, couples might also opt for a civil ceremony to save on costs and therefore the motivation would not be an aversion to religion.

## Future Research {#sec-future}
Beyond this paper, there is a lot of research that could be undertaken to further understand the presence of religion and how it influences different facets of life. Looking at marriage ceremonies specifically, more research could be done on the factors influencing why couples opt for a civil or religious ceremony. Better data collection methods could also be developed to better collect data on marriage solemnization outside of the legal form to gain insight into the prevalence of religious ceremonies that are conducted outside of legally approved premises. By collecting a greater depth of marriage ceremony data, future research will be able to better ascertain or disprove whether or not marriage ceremony data is a good reflection of religiosity in a country.

Other areas that could be examined as an indicator of a country or population's religiosity such as enrollment in religious versus secular schools, support or use of abortion services, distribution of charitable donations, and observance of religious holidays on mandated holidays for that purpose. Delving further into research on charitable donations, examining the annual charitable donations of a country and analyzing the amounts that go to each type of charity would be a novel way to see how religiosity or lack thereof can impact donation decisions.
